Parallelizing Linear Recurrent Neural Nets Over Sequence Length

### 2017-10-27 paper/main.tex
Delete comments

In order to illustrate that the linearization does not necessarily come at the
cost of expressibility, we show that the GILR-LSTM architecture, computed with
the parallel linear recurrence algorithm, is able to outperform the optimized CuDNN LSTM
implementation
on a pathological example from the original LSTM paper
\cite{hochreiter1997long}. Finally we show that our GILR-LSTM is able to
quickly achieve state of the art results on an existing medical data
set. \Chris{I hope so\ldots}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The computational performance of the LSTM and GILR-LSTM models were compared       %
% across a wide range of minibatch sizes and sequence lengths. Although the 234    %
% unit GILR-LSTM has a similar number of parameters to the 256 unit LSTM, we         %
% measured the computational performance of a 256 unit GILR-LSTM to observe any      %
% impact of the linear surrogate calculation. The serial evaluation of the LSTM    %
% causes its throughput to be independent of the sequence length as a doubling of  %
% sequence length causes a doubling of runtime. Notably, the GILR-LSTM can achieve a %
% higher throughput by running on one 8192 event sequence at a time than an LSTM   %
% running with 256 sequence minibatches. Similar speedups were found for           %
% inference.                                                                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% % Experiments were performed on the N-MNIST \cite{orchard2015converting}
% dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by
% panning % an event driven camera over each digit. Each example in N-MNIST is a
% sequence of single pixel events (x, y, polarity, timestamp) where % x and y each
% indicate a pixel position in [0, 33], polarity indicates whether the pixel was
% switching on or off, and timestamp is the time of the event in
% microseconds. Videos produced from the event data show that positive polarity
% events are often located on the leading edge of the digit motion and negative
% polarity events on the trailing edge. Between the 60,000 digits in the training
% set, N-MNIST contains approximately 250 million pixel events with sequence
% lengths ranging from 500 to 8000 and averaging 4000 events.

% % We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with
% 256 units per layer and a two layer GILR-LSTM with 234 units per layer. The
% GILR-LSTM layer size was selected so that it had slightly fewer parameters than
% the LSTM. Each model transformed the incoming event position into a
% 40-dimensional embedding vector which was then combined with the polarity to
% produce a 41 dimensional input. Both models output a $2 \mathbf{x} 34^2$ matrix
% containing two future event location probability distributions conditioned on
% the polarity of the future event. The training algorithm only considers the
% probability distribution of the true future polarity and uses the cross entropy
% loss function. The Adam \cite{kingma2014adam} optimization algorithm and Glorot
% \cite{glorot2010understanding} initialization scheme were used. Training on a
% minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling
% $b$ N-MNIST sequences and then extracting a single random $T$ element
% subsequence (and its 50 element ahead forecast) from the sequence. Sequences
% less than $T$ elements were padded out to length $T$, and there was a "burn-in"
% of 30 events at the start of each sequence where no predictions were made. An
% epoch was defined as a pass over as many pixel events as there are in the full
% training set.

% % All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow}
% on a single Nvidia K80 GPU running for up to 18 hours. The LSTM model was
% computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are
% slower than but algorithmically similar to the cuDNN LSTM implementation.

% % \subsection{Training performance} % The speed of neural net training is not
% solely determined by the training throughput but also by the frequency of
% optimization steps. As an example, a batch method may process inputs at the same
% rate as a minibatch method, but the minibatch method will generally converge
% much faster on large datasets. On an infinite dataset, the batch method never
% takes a single optimization step regardless of its throughput. With this example
% in mind, it is clear that achieving fast training is a balancing act between
% training data throughput and optimization step latency and that training
% performance should be evaluated with training curves and not just throughput
% numbers such as time per epoch.

% % The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs. LSTM
% throughput depends only on minibatch size but LSTM latency depends on both
% minibatch size and sequence length. Experiments were conducted training the LSTM
% models with minibatch size 256 and sequence lengths 128, 256, 512, and
% 1024. Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to
% faster initial learning but inferior final performance.

% % The throughput and latency of the GILR-LSTM are both influenced by batch size
% and sequence length. Several GILR-LSTMs were trained with batch size 16 and
% sequence length 1024. This combination was chosen because of the nearly maximum
% throughput, the low latency, and the ease of building a minibatch given the
% distribution of sequence lengths in the data set. The 234 unit GILR-LSTM reaches
% a better training loss in roughly 4.5 hours than any of the LSTMs could reach in
% 18 hours. This experimental evidence indicates the GILR-LSTM is as powerful of a
% model as the LSTM and can be trained in a fraction of the time.

% \begin{figure}[t]
% \centering
% \includegraphics[width=12cm]{lc2.png}
% \caption{The LSTM models had a direct relationship between sequence length and training curve: the shorter the sequence length, the faster the initial learning and
% the higher the final loss. The faster initial learning is explained by the decreased
% latency of optimization steps and the higher final loss is explained by the
% inability to learn long dependencies. The GILR-LSTM has 234 units per hidden
% layer to have the same number of parameters as the 256 unit LSTMs.}
% \label{fig:learning_curves}
% \end{figure}

% \subsection{Test performance}
% \begin{table}[h]
%   \caption{N-MNIST test results}
%   \label{test-results}
%   \centering
%   \begin{tabular}{lllll}
%     \toprule
%     & Train Cross Entropy     & Test Cross Entropy     & Top-5 Accuracy & Top-20 Accuracy \\
%     \midrule
%     GILR-LSTM 			&  \textbf{4.567} & 4.856           & 10.51\% & 36.70\%   \\
%     LSTM, seqlen=128    & 4.664 & \textbf{4.846}  & \textbf{10.62\%} & \textbf{37.08\%}   \\
%     LSTM, seqlen=256    & 4.620                & 4.906           & 10.11\% & 35.40\%   \\
%     LSTM, seqlen=512    & 4.602                & 4.883           & 10.45\% & 36.32\%   \\
%     LSTM, seqlen=1024   & 4.611                & 4.880           & 10.32\% & 35.92\%   \\
%     \bottomrule
%   \end{tabular}
% \end{table}
% Beyond cross entropy on train and test, we also evaluated top-$k$ accuracy for $k=5, 20$. A probability distribution is top-$k$ correct if it assigns the realized location one of the
% $k$ largest probabilities. N-MNIST contains 1156 pixel positions, so top-5 and top-20 accuracy are equivalent to localizing the future to 0.43\% and 1.7\% of pixels.

% No regularization was attempted and all of the models overfit, as indicated by the best test performance from the model with the worst train performance. Table \ref{test-results} contains the full test results. Our focus was the fast training of powerful models, and we leave regularizing parallel linear recurrences and LS-RNNs to future work. Although not tested, it is possible that the GILR-LSTM generalized better than the LSTMs trained on long sequences due to the much smaller minibatch size of GILR-LSTM leading to a wider minima.

>>>>

In order to illustrate that the linearization does not necessarily come at the
cost of expressibility, we show that the GILR-LSTM architecture computed with
the parallel linear recurrence algorithm, is able to outperform the optimized CuDNN LSTM
implementation
on a pathological example from the original LSTM paper
\cite{hochreiter1997long}. Finally we show that our GILR-LSTM is able to
quickly achieve state of the art results on an existing medical data
set. \Chris{I hope so\ldots}

@@

@@
